19Q2

### 2019-11-20 19Q2.Rmd
add Spearman GE/CN lineage code, did not run

# WILCOXON

@@

**10/29 NEED TO FIX CIRCLE ANNOTS ON FUSION/STEELBLUE4 HEATMAP**

@@

  print("Making heatmap object.")
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = na_color,
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                row_title = row_title_txt,
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  # if(color == "darkgreen") {
                  #   if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) { radius <- height*0.16 }
                  #   if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) { radius <- height*0.3 }
                  #   else { radius <- width*0.17 }
                  # }
                  # if(color == "steelblue4") {
                  #   if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) { radius <- height*0.5 }
                  #   if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) { radius <- height*0.16 }
                  #   else { radius <- width*0.18 }
                  # }
                  # if(color == "turquoise4") {
                  #   if(!is.na(str_match(string = tolower(feature), pattern = "fus|prot"))) { radius <- height*0.16 }
                  #   else { radius <- width*0.18 }
                  # }
                  # if(color == "darkorchid4") {
                  #   if(is.na(str_match(string = tolower(feature), pattern = "prot"))) { radius <- height*3.3 }
                  #   if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) { radius <- height*2 }
                  # }
                  # 
                  grid.text(signif_grid_annot1[i, j], x = x + width*0.25, y = y + height*0.065, gp = gpar(fontsize = 17))

@@

# SPEARMAN: PanCan

@@

## GE/CN

Spearman correlations for gene expression and copy number:
```{r spear_calc, warning=FALSE, eval=FALSE}
# CRISPR
crispr_spear_cn <- crispr %>% group_by(Gene) %>% drop_na(CN) %>%
  summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
            CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value,
            CN_p_log10 = -log10(CN_p),
            InG2P = "No",
            Dataset = "CRISPR",
            Metric = "Copy Number")
crispr_spear_cn <- transform(crispr_spear_cn, Gene = reorder(Gene, CN_r))
crispr_spear_cn <- add_column(crispr_spear_cn, Drug = "CRISPR", .before = "Gene")
Drug_Gene <- with(crispr_spear_cn, paste0(Drug, "_", Gene))
crispr_spear_cn <- add_column(crispr_spear_cn, Drug_Gene = Drug_Gene, .before = "Drug")
saveRDS(crispr_spear_cn, "./data_munging/rds/crispr_cn_cor.rds")

crispr_spear_ge <- crispr %>% group_by(Gene) %>% drop_na(TPM) %>%
  summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
            GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
            GE_p_log10 = -log10(GE_p),
            InG2P = "No",
            Dataset = "CRISPR",
            Metric = "Gene Expression")
crispr_spear_ge <- transform(crispr_spear_ge, Gene = reorder(Gene, GE_r))
crispr_spear_ge <- add_column(crispr_spear_ge, Drug = "CRISPR", .before = "Gene")
Drug_Gene <- with(crispr_spear_ge, paste0(Drug, "_", Gene))
crispr_spear_ge <- add_column(crispr_spear_ge, Drug_Gene = Drug_Gene, .before = "Drug")
saveRDS(crispr_spear_ge, "./data_munging/rds/crispr_ge_cor.rds")

# CCLE
ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
ccle_filt <- unfactorize(ccle_filt)
ccle_ge_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>%
  summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
            GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value, 
            GE_p_log10 = -log10(GE_p))
ccle_ge_res$InG2P <- ifelse(ccle_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
ccle_ge_res$Dataset <- "CCLE"

ccle_cn_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>%
  summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
            CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value,
            CN_p_log10 = -log10(CN_p))
ccle_cn_res$InG2P <- ifelse(ccle_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
ccle_cn_res$Dataset <- "CCLE"

# CTRP
ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
ctrp_filt <- unfactorize(ctrp_filt)
ctrp_ge_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>%
  summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
            GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value,
            GE_p_log10 = -log10(GE_p))
ctrp_ge_res$InG2P <- ifelse(ctrp_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
ctrp_ge_res$Dataset <- "CTRP"

ctrp_cn_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>%
  summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
            CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value,
            CN_p_log10 = -log10(CN_p))
ctrp_cn_res$InG2P <- ifelse(ctrp_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
ctrp_cn_res$Dataset <- "CTRP"

# GDSC
gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes)
gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
gdsc_filt <- unfactorize(gdsc_filt)
gdsc_ge_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>%
  summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
            GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value,
            GE_p_log10 = -log10(GE_p))
gdsc_ge_res$InG2P <- ifelse(gdsc_ge_res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])), "Yes", "No")
gdsc_ge_res$Dataset <- "GDSC"

gdsc_cn_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>%
  summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
            CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value,
            CN_p_log10 = -log10(CN_p))
gdsc_cn_res$InG2P <- ifelse(gdsc_cn_res$Drug_Gene %in% unique(as.character(g2p_amp$Drug_Gene[!is.na(g2p_amp$Drug_Gene)])), "Yes", "No")
gdsc_cn_res$Dataset <- "GDSC"

# All
cn_cor <- rbind(ccle_cn_res, ctrp_cn_res, gdsc_cn_res)
saveRDS(cn_cor, "./data_munging/rds/drug_cn_cor.rds")
ge_cor <- rbind(ccle_ge_res, ctrp_ge_res, gdsc_ge_res)
saveRDS(ge_cor, "./data_munging/rds/drug_ge_cor.rds")
```

@@

@@

# SPEARMAN: Lineages

## GE/CN

# OLD

>>>>

# WILCOXON: Mutations

@@

@@

  print("Making heatmap object.")
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = na_color,
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                row_title = row_title_txt,
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.text(signif_grid_annot1[i, j], x = x + width*0.25, y = y + height*0.065, gp = gpar(fontsize = 17))

@@

# SPEARMAN: GE and CN

@@

```{r}
MakeDrugscreenSpearmanCorrDF <- function(dataset, df, metric, lineage) {
  df_filt <- filter(df, Drug %in% g2p_drugs & Gene %in% g2p_genes)
  df_filt$Gene <- factor(df_filt$Gene, ordered = FALSE)
  df_filt <- unfactorize(df_filt)

  if(metric == "GE") {
    if(lineage) {
      res <- df_filt %>% group_by(Drug, Gene, disease, Drug_Gene_Lin) %>% drop_na(TPM) %>%
        summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
                  GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value, 
                  GE_p_log10 = -log10(GE_p))
      res$InG2P <- ifelse(res$Drug_Gene_Lin %in% unique(as.character(g2p_expr$Drug_Gene_Lin[!is.na(g2p_expr$Drug_Gene_Lin)])),   "Yes", "No")
    }
    if(!lineage) {
      res <- df_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>%
        summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
                  GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value, 
                  GE_p_log10 = -log10(GE_p))
      res$InG2P <- ifelse(res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])),   "Yes", "No")
    }
  }
  if(metric == "CN") {
    if(lineage) {
      res <- df_filt %>% group_by(Drug, Gene, disease, Drug_Gene_Lin) %>% drop_na(CN) %>%
        summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
                  CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value,
                  CN_p_log10 = -log10(CN_p))
      res$InG2P <- ifelse(res$Drug_Gene_Lin %in% unique(as.character(g2p_amp$Drug_Gene_Lin[!is.na(g2p_amp$Drug_Gene_Lin)])), "Yes", "No")
    }
    if(!lineage) {
      res <- df_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>%
        summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
                  CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value, 
                  CN_p_log10 = -log10(GE_p))
      res$InG2P <- ifelse(res$Drug_Gene %in% unique(as.character(g2p_expr$Drug_Gene[!is.na(g2p_expr$Drug_Gene)])),   "Yes", "No")
    }
  }
  res$Dataset <- dataset
  return(res)
}
```

## Lineage-agnostic

```{r spear_calc, warning=FALSE, eval=FALSE}
# CRISPR
crispr_spear_cn <- crispr %>% group_by(Gene, disease) %>% drop_na(CN) %>%
  summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
            CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value,
            CN_p_log10 = -log10(CN_p),
            InG2P = "No",
            Dataset = "CRISPR",
            Metric = "Copy Number")
crispr_spear_cn <- transform(crispr_spear_cn, Gene = reorder(Gene, CN_r))
crispr_spear_cn <- add_column(crispr_spear_cn, Drug = "CRISPR", .before = "Gene")
Drug_Gene <- with(crispr_spear_cn, paste0(Drug, "_", Gene))
crispr_spear_cn <- add_column(crispr_spear_cn, Drug_Gene = Drug_Gene, .before = "Drug")
crispr_spear_cn <- add_column(crispr_spear_cn, Drug_Gene_Lin = paste0(Drug_Gene, "_", disease), .before = "Drug")
saveRDS(crispr_spear_cn, "./data_munging/rds/crispr_cn_cor_lineage.rds")

crispr_spear_ge <- crispr %>% group_by(Gene, disease) %>% drop_na(TPM) %>%
  summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
            GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
            GE_p_log10 = -log10(GE_p),
            InG2P = "No",
            Dataset = "CRISPR",
            Metric = "Gene Expression")
crispr_spear_ge <- transform(crispr_spear_ge, Gene = reorder(Gene, GE_r))
crispr_spear_ge <- add_column(crispr_spear_ge, Drug = "CRISPR", .before = "Gene")
Drug_Gene <- with(crispr_spear_ge, paste0(Drug, "_", Gene))
crispr_spear_ge <- add_column(crispr_spear_ge, Drug_Gene = Drug_Gene, .before = "Drug")
crispr_spear_ge <- add_column(crispr_spear_ge, Drug_Gene_Lin = paste0(Drug_Gene, "_", disease), .before = "Drug")
saveRDS(crispr_spear_ge, "./data_munging/rds/crispr_ge_cor_lineage.rds")

ccle_cn_res <- MakeSpearmanCorrDF(dataset = "CCLE", df = ccle, metric = "CN", lineage = FALSE)
ctrp_cn_res <- MakeSpearmanCorrDF(dataset = "CTRP", df = ctrp, metric = "CN", lineage = FALSE)
gdsc_cn_res <- MakeSpearmanCorrDF(dataset = "GDSC", df = gdsc, metric = "CN", lineage = FALSE)
cn_cor <- rbind(ccle_cn_res, ctrp_cn_res, gdsc_cn_res)
saveRDS(cn_cor, "./data_munging/rds/drug_cn_cor.rds")

ccle_ge_res <- MakeSpearmanCorrDF(dataset = "CCLE", df = ccle, metric = "GE", lineage = FALSE)
ctrp_ge_res <- MakeSpearmanCorrDF(dataset = "CTRP", df = ctrp, metric = "GE", lineage = FALSE)
gdsc_ge_res <- MakeSpearmanCorrDF(dataset = "GDSC", df = gdsc, metric = "GE", lineage = FALSE)
ge_cor <- rbind(ccle_ge_res, ctrp_ge_res, gdsc_ge_res)
saveRDS(ge_cor, "./data_munging/rds/drug_ge_cor.rds")
```

@@

## Lineage

## Lineage

```{r spear_calc_lin, eval=FALSE}
# CRISPR
# crispr_spear_cn <- crispr %>% group_by(Gene) %>% drop_na(CN) %>%
#   summarize(CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
#             CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value,
#             CN_p_log10 = -log10(CN_p),
#             InG2P = "No",
#             Dataset = "CRISPR",
#             Metric = "Copy Number")
# crispr_spear_cn <- transform(crispr_spear_cn, Gene = reorder(Gene, CN_r))
# crispr_spear_cn <- add_column(crispr_spear_cn, Drug = "CRISPR", .before = "Gene")
# Drug_Gene <- with(crispr_spear_cn, paste0(Drug, "_", Gene))
# crispr_spear_cn <- add_column(crispr_spear_cn, Drug_Gene = Drug_Gene, .before = "Drug")
# saveRDS(crispr_spear_cn, "./data_munging/rds/crispr_cn_cor.rds")
# 
# crispr_spear_ge <- crispr %>% group_by(Gene) %>% drop_na(TPM) %>%
#   summarize(GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
#             GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
#             GE_p_log10 = -log10(GE_p),
#             InG2P = "No",
#             Dataset = "CRISPR",
#             Metric = "Gene Expression")
# crispr_spear_ge <- transform(crispr_spear_ge, Gene = reorder(Gene, GE_r))
# crispr_spear_ge <- add_column(crispr_spear_ge, Drug = "CRISPR", .before = "Gene")
# Drug_Gene <- with(crispr_spear_ge, paste0(Drug, "_", Gene))
# crispr_spear_ge <- add_column(crispr_spear_ge, Drug_Gene = Drug_Gene, .before = "Drug")
# saveRDS(crispr_spear_ge, "./data_munging/rds/crispr_ge_cor.rds")

ccle_cn_res_lin <- MakeSpearmanCorrDF(dataset = "CCLE", df = ccle, metric = "CN", lineage = TRUE)
ctrp_cn_res_lin <- MakeSpearmanCorrDF(dataset = "CTRP", df = ctrp, metric = "CN", lineage = TRUE)
gdsc_cn_res_lin <- MakeSpearmanCorrDF(dataset = "GDSC", df = gdsc, metric = "CN", lineage = TRUE)
cn_cor_lin <- rbind(ccle_cn_res_lin, ctrp_cn_res_lin, gdsc_cn_res_lin)
saveRDS(cn_cor_lin, "./data_munging/rds/drug_cn_cor_lineage.rds")

ccle_ge_res_lin <- MakeSpearmanCorrDF(dataset = "CCLE", df = ccle, metric = "GE", lineage = TRUE)
ctrp_ge_res_lin <- MakeSpearmanCorrDF(dataset = "CTRP", df = ctrp, metric = "GE", lineage = TRUE)
gdsc_ge_res_lin <- MakeSpearmanCorrDF(dataset = "GDSC", df = gdsc, metric = "GE", lineage = TRUE)
ge_cor_lin <- rbind(ccle_ge_res_lin, ctrp_ge_res_lin, gdsc_ge_res_lin)
saveRDS(ge_cor_lin, "./data_munging/rds/drug_ge_cor_lineage.rds")
```

@@

# OLD CODE
